Remote sensing riverine gas exchange and carbon efflux in ungauged rivers from SWOT observations

### 2021-06-04 manuscript_v3.Rmd
some ms updates

##Highlights (3-5 points, 85 characters each w/ spaces)
- The BIKER algorithm indirectly remotely senses riverine gas exchange velocity
- Predicted gas exchange velocities compare well with in-situ methods
- BIKER predicts $CO_2$ fluxes when coupled with in situ gas concentration data
- Expected SWOT measurement errors exert a marginal influence on BIKER's estimates
- Errors are dominated by process-level uncertainties, not parameter inversion

@@

##Abstract
Go BIKER!!
*Note: this is for Remote Sensing of Environment (I think). So 400 words here.*

##1 Introduction
**I don't like this opening paragraph** Vital to sustaining life, the Earth's carbon cycle moves carbon between the atmosphere, geosphere, hydrosphere, and biosphere via a complex set of interconnected biogeochemical processes. The primary connection between the terrestrial and oceanic biospheres is the global river network, which exports much carbon from the continents to the oceans [@colePlumbingGlobalCarbon2007a]. However en route to the oceans, this river network additionally transports and transforms sediments, nutrients, organic matter, dissolved gases, and other constituents. Organic matter and inorganic nutrients are also produced and/or consumed within rivers via ecosystem metabolism [@odumPrimaryProductionFlowing1956]. One net effect of all of this is the global river network is usually supersaturated with dissolved gases and evades substantial amounts of carbon from its water surface to the atmosphere [@colePlumbingGlobalCarbon2007a].

Current estimates of total carbon dioxide ($CO_2$) evasion to the atmosphere vary from 650-1800 Tg C/yr [@lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013a], with 167 Tg-C/yr coming from mountain streams alone [@horgbyUnexpectedLargeEvasion2019]. Further, in-river metabolism is increasingly understood to contribute to this flux [@hotchkissSourcesProcessesControlling2015]. Equation 1 represents this flux for any sparingly soluble gas $[M/L^2T]$ given the gas concentration gradient between the water $C_{water} [M/L^3]$ and the air $C_{air} [M/L^3]$ and the gas exchange velocity $k [L/T]$. Given the structure of equation 1, calculations of this flux are very sensitive to one's measurements/estimates of *k*. Note that *k* is different for different gases and water temperatures [see @hallGasExchangeStreams2020 for a thorough review of gas exchange in rivers].

@@

*k* is also vital to understanding freshwater ecology through bulk river metabolism [more specifically, gross primary production and ecological respiration- @bernhardtMetabolicRegimesFlowing2018]. The general mass-balance equation for river metabolism (*M* $[M/L^2T]$) is provided as equation 2, where *DO* is the dissolved oxygen concentration $[M/L^3]$, *D* is river depth [L], and $k_{o_2}$ is *k* for oxygen at some temperature. When viewing equations 1 and 2 in aggregate, it is clear that *k* plays a critical role in two of the core, fundamental biogeochemical processes that occur in river networks. Both are reliant on accurate and high-fidelity assessments of *k*, as well as a rigorous understanding of the physical mechanisms driving *k*.
$$\mathbf{(2)} \frac{dDO}{dt}=\frac{M}{D}+\frac{k_{o_2}}{D}(DO_{air}-DO_{water})$$
Researchers use a range of methods to measure *k*, from floating domes to eddy covariance towers to measure at a specific point in the river. They also inject tracer gases to estimate reach-scale gas exchange [@hallGasExchangeStreams2020]. These approaches are ideal but are infeasible when working at the river network scale across potentially tens of thousands of rivers. So, predictive models are used to estimate *k* from readily available river geomorphology [e.g. @oconnorMechanismReaerationNatural1958; @palumboAssessingPerformanceReaeration2014; @raymondScalingGasTransfer2012a]. These models often, though not always, predict $k_{600}$ or *k* that has been normalized to a Schmidt number of 600 ($CO_2$ at 20 degrees Celsius or $O_2$ at 17.5 degrees).

@@

Many of these predictive models are based on the observation that *k* is generally correlated with both the turbulent energy dissipation rate *eD* $[L^2/T^3]$ [@ulsethDistinctAirWater2019a; @raymondScalingGasTransfer2012a] and the water surface surface shear velocity $U_\star [L/T]$ [@wangPhysicallyBasedScaling2021; @katulStructureFunctionModel2018]. This is convenient because both *eD* and $U_\star$ are calculated using easily estimated hydraulic parameters: $eD \approx gS_hV$ [@tsivoglouTracerMeasurementReaeration1976] and $U_\star=\sqrt{gS_hR_h}$, where *g* is gravitational acceleration $[L/T^2]$, $S_h$ is water surface slope $[L/L]$, *V* is average flow velocity $[L/T]$ and $R_h$ is the hydraulic radius [L] (which often approximates *D*). Because slope is readily available in any hydrographic data product, most efforts to predict *k* across thousands of rivers are thereby limited by the quality of the final parameters *V* or $R_h$. Since both must be measured in the field or themselves estimated by empirical models, uncertainty in *V* or $R_h$ dominates errors in *k* estimates. This is exacerbated in ungauged basins that cover large areas, especially in the carbon-rich Arctic inland waters, as no in situ hydraulic information is available and fieldwork is impractical [@gleasonRemoteSensingRiver2020].

@@

For the purposes of predicting *k* from *V* or $R_h$, biogeochemists have further extended HG models by training them on large hydraulics datasets for thousands of rivers and under all streamflow conditions [e.g @borgesGloballySignificantGreenhousegas2015a; @horgbyUnexpectedLargeEvasion2019; @lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013a]. However, it has long been established that HG parameters are either cross-section specific and highly variable from river to river (at-a-station HG) or they are discharge specific [downstream HG- @gleasonHydraulicGeometryNatural2015; @parkWorldwideVariationsHydraulic1977; @rhodesBfmDiagramGraphical1977]. This suggests that the 'global-scope' HG models that are usually used are also sensitive to their training data and can potentially yield pretty different hydraulics for the same streamflow. It is currently not well understood how sensitive global estimates of riverine gas evasion are to the specific 'global-scope' HG model that is employed by the worker.

@@

To answer these questions, we developed a new McFLI algorithm that ingests SWOT data and produces $k_{o_2}$ estimates and their explicit Bayesian uncertainities that requires no in situ inputs of any kind (although in situ data can be ingested and will improve results). We name the RS of $k_{o_2}$ algorithm BIKER, or the 'Bayesian Inference/Inversion of the $k_{o_2}$ Evasion Rate'. We validate BIKER on simulated SWOT data for 49 SWOT-observable rivers from around the world using hydraulic models to produce SWOT-like data as SWOT has not yet launched. This is standard practice in the SWOT community [@frassonExploringFactorsControlling2021; @durandIntercomparisonRemoteSensing2016b]. We also quantify BIKER's sensitivity to the expected SWOT measurement errors on 17 of those rivers: while SWOT data represent a sea change in inland water monitoring, it is expected to have an approximately 10cm error in water surface elevation [@biancamariaSWOTMissionIts2016], as well as river width errors [@frassonExploringFactorsControlling2021]. Finally, we use previously published dissolved $CO_2$ data to represent a hypotetical in situ $CO_2$ sensor and compare the bulk carbon efflux from the 49 rivers as calculated using BIKER and previously published in situ techniques.

##2 Data
Numerous datasets were used in this study to develop and validate BIKER and are detailed below.

BIKER validation (section 3.3) was performed on 49 SWOT-simulated rivers. Because SWOT has yet to launch, it is standard practice to benchmark SWOT-related algorithms on SWOT-like data. There are three current types of SWOT-like data: 1) AirSWOT, which is an airborne Ka-band inSAR currently limited to five rivers globally, 2) simulated rivers that mimic the type of data SWOT will provide, and 3) the SWOT simulator, which introduces measurement errors to these simulated rivers. Because we are principally interested in algorithm performance, we limit our validation setup to simulated rivers in order to benchmark across as many rivers as possible. These simulated rivers are simply reach-averaged hydraulic model outputs where the water surface heights and widths are labelled as RS observations and are used as the sole inputs to BIKER. Here, we use 49/51 rivers collected by @frassonExploringFactorsControlling2021 and @durandIntercomparisonRemoteSensing2016b. These are the two benchmarking studies that have explored RSQ algorithm performance for the SWOT mission. Please consult both of those papers for all of the rivers' locations as well as hydraulic model specifications. Ultimately, the 49 rivers are spread across the United States, France, Italy, the United Kingdom, and Bangladesh. We omit both models for the Saint Lawrence River, Canada from @durandIntercomparisonRemoteSensing2016b because they lack enough hydraulic information to calculate $U_\star$.

@@

BIKER prior specifications (section 3.2.2) require the use of a training dataset of field-measured river hydraulics. To do this, we use a previously published compilation of field measurements that were originally made to calibrate United States Geological Survey (USGS) streamgauge rating curves [@brinkerhoffReconcilingAtaStationAtManyStations2019]. That dataset contains over 530,000 unique measurements of average channel velocity, width, depth, and discharge from across the continental United States.

For the $CO_2$ evasion and carbon efflux calculations (section 3.4), we use 26 bi-weekly dissolved $CO_2$ samples made by @beaulieuControlsGasTransfer2012a at one location in the Ohio River for one calender year from 2008-2009 (Figure S1). Note that this $CO_2$ data is for the Ohio River only but was applied to all 49 rivers (which includes multiple sections of the Ohio River). Because we are exclusively interested in the relative differences between $FCO_2$ estimates and not the raw fluxes themselves, any $CO_2$ data representative of SWOT-observable rivers was deemed acceptable. **Kelly just published hourly CO2 data for the CT, maybe we should use that instead to get mean daily values?**

Finally, we also made extensive use of data and models developed by workers to predict gas evasion rates from river hydraulics. Specifically, we used the data measured, collected, or simulated by @raymondScalingGasTransfer2012a, @ulsethDistinctAirWater2019a, and @wangPhysicallyBasedScaling2021.

##3 Methods
To build BIKER, we join a process-based model for $k_{o_2}$ (section 3.1) with a McFLI framework for inverting SWOT measurements via Bayesian inference (section 3.2). Following the description of that process, we describe the validation setup (section 3.3) and the workflow for comparing estimated bulk carbon effluxes from a suite of 'global-scope' HG models (section 3.4). A flowchart detailing the entire study is provided as Figure 1.

![Figure 1. Flowchart of methodology used in this study. We begin by choosing a process-based model for k (section 3.1) that is a function of river hydraulics measurable by SWOT. Then, we implement this model within a McFLI ('BIKER') to estimate k solely from river width and water surface slope (section 3.2). Finally, we couple BIKER with in situ CO2 data to compare BIKER-estimated gas fluxes with established in situ methods (section 3.4). See section 3.3 for the validation setup.](cache/flowchart.jpg)

###3.1 Choosing a predictive model for *k*
To predict *k* from just the SWOT observables, a physical model for *k* is first selected. As mentioned in section 1, there are nearly two dozen predictive models for *k* that have been devolped since the 1950s [@wangPhysicallyBasedScaling2021; @hallGasExchangeStreams2020]. Recently, many of these models were reexplored in rivers and streams by @wangPhysicallyBasedScaling2021, who significantly expanded the existing training datasets of field-measured $k_{o_2}$ by using the *streamMetabolizer* model to calibrate a form of equation 2 to high-fidelity in situ DO datasets [@applingOvercomingEquifinalityLeveraging2018] at 35 rivers across the United States. They showed that two $k_{o_2}$ models deduced from classic, process-based theories for mass transport yield approximately identical parameters when fit to ethier field measurements or this new dataset of simulated $k_{o_2}$. They also showed through cross-validation that these models are robust to overfitting on specific sets of data (a concern with many of the aforementioned predictive models- section 1). This suggests that there are uniform scaling realtionships between certain hydraulic properties and $k_{o_2}$ that exist regardless of the training data used. @wangPhysicallyBasedScaling2021's better fit, and more parsimonious, of the two models is reprinted as equation 3 (with that paper's reported coefficients of determination $r^2$). $\alpha$ is the fitted parameter in the linear regression. Equation 3-2nd line was fit to 588 field measurements of $k_{o_2}$ while equation 3-third line was fit to 3,919 simulated $k_{o_2}$ values at 35 rivers.
$$\mathbf{(3)} \begin{eqnarray} k_{o_2} &=& \alpha U_\star \\ &=& 47.67U_\star (r^2:0.53, measured) \\ &=& 49.71U_\star (r^2: 0.76, simulated)  \end{eqnarray}$$

We implement this model within BIKER (using 48 as a reasonable value for $\alpha$ given both forms of equation 3) for a few reasons. First, @wangPhysicallyBasedScaling2021 provides compelling evidence that an $\alpha$ of approximately 48 is uniform across rivers and streams, regardless of the training data used. This is encouraging for use within BIKER as BIKER is specifically designed to be as river-agnostic as possible and so we sought as generalized an equation as possible. Second, it yields a simple linear relationship across all rivers, regardless of their size or steepeness. This is not necessairly true of other *k* models, for example those based on *eD* (section 1). Recent work has shown that *k* does not scale with *eD* via the same statistical parameters across all river geomorphologies [@ulsethDistinctAirWater2019a]. We further explored the relationship presented in that paper and found that the $k \sim eD$ relationship fundamentally breaks down in rivers with low *eD*, which corresponds to nearly all rivers that SWOT will observe (**Figure S2, need to do**). Therefore, an *eD* based predictive model is less useful in SWOT rivers. Third, we sought to explicitly predict $k_{o_2}$, as its use extends beyond constraining gas fluxes from rivers and into parameterizing river metabolism models (equation 2). We stress that most other predictive models for *k* could be implemented within BIKER. However, our goal of using BIKER on any river that is observable by SWOT necessitates that we pick an equation that is applicable across as many rivers as possible. Future work should look at using a river-specific *k* equation to pseudo-calibrate BIKER to a specific river.

@@

BIKER, and Bayesian inference in general, starts from Bayes rule (equation 4). In equation 4, $\Theta$ is some set of non-remotely-sensible parameters we want to solve for (including $k_{o_2}$), *x* is the observed data, $f(x|\Theta)$ is the 'likelihood function' or sampling model conditional on the parameters, and $p(\Theta)$ is the joint prior distribution of the parameters. Therefore, we are interested in solving for $p(\Theta|x)$, or the 'posterior' distribution. For BIKER, *x* is the SWOT-observables *W* and *H*. Note that $p(x)$ is usually computationally intractable to integrate, so Bayesian inference tools require only the proportionality to be specified: $p(\Theta|x) \propto f(x|\Theta)p(x)$. Sampling algorithms are then used to approximate the posterior distribution, as is done in BIKER and detailed below. First, we detail our likelihood specification and then our prior specifications.

@@

####3.2.1 BIKER sampling model
To concieve of equation 3 as a likelihood function, $k_{o_2}$ first needs to be written as a function of SWOT-observables *W* and *H* (which provides water surface slope $S_h$). This algebra is carried out and results in equation 5, where we assume that $R_h \approx \frac{A}{W}$ where *A* is wetted channel area. *A* is further split into the SWOT-observable portion *dA* and the unobservable portion $A_0$ following @durandEstimatingReachaveragedDischarge2014a. *dA* is estimated assuming a rectangular river channel so that $dA \approx \delta H*W$.

@@

Next, equation 5 is written as a Bayesian sampling model to have all of the SWOT observations on the left-hand side sampled from the unknown model parameters ($A_0$ and $k_{o_2}$) on the right-hand side. This is equation 6. The $\sigma_{k_{o2}}^2$ parameter refers to the uncertainty inherent in equation 5's estimates. This will be explained in detail below. Note that in the formal model specification, equation 6 is written as a normal distribution of its log-transformed quantities.

@@

####3.2.2 Prior Specifications
Equation 6 necessitates that we specify prior distributions for parameters $A_0$ and $k_{o_2}$. Prior distributions formalize the a priori estimates (and uncertainties) for the non-remotely-sensed parameters. More intuitively, BIKER priors represent our 'prior river knowledge' of what $A_0$ and $k_{o_2}$ probably are for some river since they cannot be remotely sensed. All priors are formalized within the model as truncated normal distributions of the log-transformed terms such that $log(X) \sim N(\mu, \sigma^2)$ for $\lambda < log(X) < \epsilon$, using prior hyperparameters mean ($\mu$), standard deviation ($\sigma$), and upper ($\epsilon$) and lower bounds ($\lambda$) for any parameter *X*. In order to avoid relying on in situ information, as is the goal of BIKER and SWOT, we assign prior hyperparameters using SWOT data only. $A_0$ prior hyperparameters were assigned following the method developed by @brinkerhoffConstrainingRemoteRiver2020a They developed a set of river channel prior hyperparameters for McFLI algorithms that are entirely RS-able and reflect differential channel hydraulics as a function of river geomorphology. They used an extensive database of field measurements and machine learning to identify patterns that associate river width with the hydraulic priors needed to run McFLIs so that prior hyperparameters may be assigned to rivers using only the existing remotely sensed data. This leaves the $k_{o_2}$ hyperparameters to be defined. We assigned those using a simple slope regression model trained on an available subset of the @raymondScalingGasTransfer2012a dataset of field-measured $k_{o_2}$: equation 7 assigns $\mu$ while $\sigma^2$ was set to equation 7's log-transformed standard error: 0.77. $\lambda$ and $\epsilon$ were set to log(0.0001) and log(60), respectively.

>>>>

##Highlights
- The BIKER algorithm indirectly remotely senses riverine gas exchange velocity
- Predicted gas exchange velocities compare well with in-situ methods
- BIKER predicts $CO_2$ fluxes when coupled with in situ gas concentration data
- Expected SWOT measurement errors exert a marginal influence on BIKER's estimates
- Errors are dominated by process-level uncertainties, not parameter inversion

@@

##Abstract
Go BIKER!!
*Note: this is for Remote Sensing of Environment (I think). So 400 words here. also for highlights: (3-5 points, 85 characters each w/ spaces)*

##1 Introduction
**I don't like this opening paragraph** The Earth's carbon cycle moves carbon between the atmosphere, geosphere, hydrosphere, and biosphere via a complex set of interconnected biogeochemical processes. The primary connection between the terrestrial and oceanic biospheres is the global river network, which exports a significant amount of carbon from the continents to the oceans [@colePlumbingGlobalCarbon2007a], while additionally transporting and transforming sediments, nutrients, organic matter, dissolved gases, and other constituents. Organic matter and inorganic nutrients are also produced and/or consumed within rivers via ecosystem metabolism [@odumPrimaryProductionFlowing1956]. One net effect of all of this is the global river network is usually supersaturated with dissolved gases and evades substantial amounts of carbon from its water surface to the atmosphere [@colePlumbingGlobalCarbon2007a].

Current estimates of total carbon dioxide ($CO_2$) evasion to the atmosphere vary from 650-1800 Tg C/yr [@lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013a], with 167 Tg-C/yr coming from mountain streams alone [@horgbyUnexpectedLargeEvasion2019]. Further, in-river metabolism is increasingly understood to contribute to this flux [@hotchkissSourcesProcessesControlling2015]. Equation 1 represents this flux for any sparingly soluble gas $[M/L^2T]$ given the gas concentration gradient between the water $C_{water}$ $[M/L^3]$ and the air $C_{air}$ $[M/L^3]$ and the gas exchange velocity *k [L/T]*. Given the structure of equation 1, calculations of this flux are very sensitive to one's measurements/estimates of *k*. Note that *k* is different for different gases and water temperatures [see @hallGasExchangeStreams2020 for a thorough review of gas exchange in rivers].

@@

*k* is also vital to understanding freshwater ecology through bulk river metabolism [more specifically, gross primary production and ecological respiration- @bernhardtMetabolicRegimesFlowing2018]. The general mass-balance equation for river metabolism (*M* $[M/L^2T]$) is provided as equation 2, where *DO* is the dissolved oxygen concentration $[M/L^3]$, *D* is river depth [L], and $k_{o_2}$ is *k* for oxygen at some temperature. When viewing equations 1 and 2 in aggregate, it is clear that *k* plays a critical role in two of the core, fundamental biogeochemical processes that occur in river networks. Both are reliant on accurate and high-fidelity assessments of *k*, as well as a rigorous understanding of the physical mechanisms driving *k*.
$$\mathbf{(2)} \frac{dDO}{dt}=\frac{M}{D}+\frac{k_{o_2}}{D}(DO_{air}-DO_{water})$$
Researchers use a range of methods to measure *k*, from floating domes to eddy covariance towers to measure at a specific point in the river. They also inject tracer gases to estimate reach-scale gas exchange [@hallGasExchangeStreams2020]. These approaches are ideal but are infeasible when working in ungauged rivers or at the river network scale across thousands of rivers. So, predictive models are used to estimate *k* from readily available river geomorphology [e.g. @oconnorMechanismReaerationNatural1958; @palumboAssessingPerformanceReaeration2014; @raymondScalingGasTransfer2012a]. These models often, though not always, predict $k_{600}$ or *k* that has been normalized to a Schmidt number of 600 ($CO_2$ at 20 degrees Celsius or $O_2$ at 17.5 degrees).

@@

Many of these predictive models rely on the correlation between *k* both the turbulent energy dissipation rate *eD* $[L^2/T^3]$ [@ulsethDistinctAirWater2019a; @raymondScalingGasTransfer2012a] and the water surface surface shear velocity $U_\star [L/T]$ [@wangPhysicallyBasedScaling2021; @katulStructureFunctionModel2018]. This is convenient because both *eD* and $U_\star$ are calculated using easily estimated hydraulic parameters: $eD \approx gS_hV$ [@tsivoglouTracerMeasurementReaeration1976] and $U_\star=\sqrt{gS_hR_h}$, where *g* is gravitational acceleration $[L/T^2]$, $S_h$ is water surface slope $[L/L]$, *V* is average flow velocity $[L/T]$ and $R_h$ is the hydraulic radius [L] (which often approximates *D*). Because slope is readily available in any hydrographic data product, most efforts to predict *k* across thousands of rivers are thereby limited by the quality of the final parameters *V* or $R_h$. Since both must be measured in the field or themselves estimated by empirical models, uncertainty in *V* or $R_h$ dominates errors in *k* estimates. This is exacerbated in ungauged basins that cover large areas, especially in the carbon-rich Arctic inland waters, as little in situ hydraulic information is available and fieldwork is impractical [@gleasonRemoteSensingRiver2020].

@@

For the purposes of predicting *k* from *V* or $R_h$, biogeochemists have further extended HG models by training them on large hydraulics datasets for thousands of rivers and under all streamflow conditions [e.g @borgesGloballySignificantGreenhousegas2015a; @horgbyUnexpectedLargeEvasion2019; @lauerwaldSpatialPatternsCO22015; @raymondGlobalCarbonDioxide2013a]. However, it has long been established that HG parameters are either cross-section specific and highly variable from river to river as in at-a-station HG or they are discharge specific as in downstream HG [@gleasonHydraulicGeometryNatural2015; @parkWorldwideVariationsHydraulic1977; @rhodesBfmDiagramGraphical1977]. This suggests that the 'global-scope' HG models that are usually used are also sensitive to their training data and can potentially yield pretty different hydraulics for the same streamflow. It is currently not well understood how sensitive global estimates of riverine gas evasion are to the specific 'global-scope' HG model that is employed by the worker. Further, these approaches rely on ethier in situ discharge records or modeled streamflow which introduces additional uncertainity.

@@

To answer these questions, we developed a new McFLI algorithm that ingests SWOT data and produces $k_{o_2}$ estimates and their explicit Bayesian uncertainities that requires no in situ inputs of any kind (although in situ data can be ingested and will improve results). We name the RS of $k_{o_2}$ algorithm BIKER, or the 'Bayesian Inference/Inversion of the $k_{o_2}$ Evasion Rate'. We validate BIKER on simulated SWOT data for 49 SWOT-observable rivers from around the world using hydraulic models to produce SWOT-like data as SWOT has not yet launched. This is standard practice in the SWOT community [@frassonExploringFactorsControlling2021; @durandIntercomparisonRemoteSensing2016b]. We also quantify BIKER's sensitivity to the expected SWOT measurement errors on 17 of those rivers: while SWOT data represent a sea change in inland water monitoring, it is expected to have an approximately 10cm error in water surface elevation [@biancamariaSWOTMissionIts2016], as well as river width errors [@frassonExploringFactorsControlling2021]. Finally, we use previously published dissolved $CO_2$ data to represent a hypotetical in situ $CO_2$ sensor and compare the bulk carbon efflux from the 49 rivers as calculated using BIKER and previously published in situ techniques for predicting $k_{CO_2}$.

##2 Data
Numerous datasets were used in this study to develop and validate BIKER and are detailed below. Please see Figure 1 for a map of the approximate locations for the data used in this study.

![Figure 1: Map of the 49 hydraulic models and 1 timeseries of CO2 samples used in this study. Note that hydraulic model locations are approximate as most of the models are not geo-referenced. We additionally used over 530,000 discrete measurements of river channel hydraulics from across the continetal United States (Brinkerhoff et al. 2019) that are not mapped here.](cache/map_fin.jpg)

BIKER validation (section 3.3) was performed on 49 SWOT-simulated rivers. Because SWOT has yet to launch, it is standard practice to benchmark SWOT-related algorithms on SWOT-like data. There are three current types of SWOT-like data: 1) AirSWOT, which is an airborne Ka-band inSAR currently limited to five rivers globally, 2) simulated rivers that mimic the type of data SWOT will provide, and 3) the SWOT simulator, which introduces measurement errors to these simulated rivers. Because we are principally interested in algorithm performance, we limit our validation setup to simulated rivers in order to benchmark across as many rivers as possible. These simulated rivers are simply reach-averaged hydraulic model outputs where the water surface heights and widths are labelled as RS observations and are used as the sole inputs to BIKER. Here, we use 49/51 rivers collected by @frassonExploringFactorsControlling2021 and @durandIntercomparisonRemoteSensing2016b. These are the two benchmarking studies that have explored RSQ algorithm performance for the SWOT mission. The approximate locations of these rivers are plotted in FIgure 1, and please consult both of those papers for all of the hydraulic model specifications. Ultimately, the 49 rivers are spread across the United States, France, Italy, the United Kingdom, and Bangladesh. We omit both models for the Saint Lawrence River, Canada from @durandIntercomparisonRemoteSensing2016b because they lack enough hydraulic information to calculate $U_\star$, which will be necessary in section 3.2.

@@

BIKER prior specifications (section 3.2.2) require the use of a training dataset of field-measured river hydraulics. To do this, we use a previously published compilation of field measurements that were originally made to calibrate United States Geological Survey (USGS) streamgauge rating curves [@brinkerhoffReconcilingAtaStationAtManyStations2019]. That dataset contains over 530,000 unique measurements of average channel velocity, width, depth, and discharge from across the continental United States. It has been previously used to inform prior specifications for RSQ algorithms [@andreadisConstrainingAssimilationSWOT2020d; @brinkerhoffConstrainingRemoteRiver2020a]. We acknowledge that our priors will therefore be biased towards the types of rivers in the continental United States and miss key hydroclimatic regions like Arctic and Tropical rivers. However, previous success in using these data in unrepresented regions [@brinkerhoffConstrainingRemoteRiver2020a] bodes well for their use here. Further, 14/19 rivers are from outside of this region (Figure 1).

For the $CO_2$ evasion and carbon efflux calculations (section 3.4), we use 26 bi-weekly dissolved $CO_2$ samples made by @beaulieuControlsGasTransfer2012a at one location in the Ohio River for one calender year from 2008-2009 (Figure S1). Note that this $CO_2$ data is for the Ohio River only but was applied to all 49 rivers (which includes multiple sections of the Ohio River). Because we are exclusively interested in the relative differences between $FCO_2$ estimates and not the raw fluxes themselves, any $CO_2$ data representative of SWOT-observable rivers was deemed acceptable.

Finally, we also made extensive use of data and models developed by workers to predict gas evasion rates from river hydraulics. Specifically, we used the data measured, collected, or simulated by @raymondScalingGasTransfer2012a, @ulsethDistinctAirWater2019a, and @wangPhysicallyBasedScaling2021. Note that much of this data was collated from many other datasets from many other authors.

##3 Methods
To build BIKER, we join a predictive model for $k_{o_2}$ (section 3.1) with a McFLI framework for inverting SWOT measurements via Bayesian inference (section 3.2). Following the description of that process, we describe the validation setup (section 3.3) and the workflow for comparing estimated bulk carbon effluxes from a suite of 'global-scope' HG models (section 3.4). A flowchart detailing the entire study is provided as Figure 2.

![Figure 2. Flowchart of methodology used in this study. We begin by choosing a process-based model for k (section 3.1) that is a function of river hydraulics measurable by SWOT. Then, we implement this model within a McFLI ('BIKER') to estimate k solely from river width and water surface slope (section 3.2). Finally, we couple BIKER with in situ CO2 data to compare BIKER-estimated gas fluxes with established in situ methods (section 3.4). See section 3.3 for the validation setup.](cache/flowchart.jpg)

###3.1 Choosing a predictive model for *k*
To predict *k* from just the SWOT observables, a physical model for *k* is first selected. As mentioned in section 1, there are dozens of predictive models for *k* that have been devolped since the 1950s [@hallGasExchangeStreams2020]. Recently, many of these models were re-explored in rivers and streams by @wangPhysicallyBasedScaling2021, who significantly expanded the existing training datasets of field-measured $k_{o_2}$ by using the *streamMetabolizer* model to calibrate a form of equation 2 to high-fidelity in situ DO datasets [@applingOvercomingEquifinalityLeveraging2018] at 35 rivers across the United States. They showed that two $k_{o_2}$ models deduced from classic, process-based theories for mass transport yield approximately identical parameters when fit to ethier field measurements or this new dataset of simulated $k_{o_2}$. They also showed through cross-validation that these models are more robust to overfitting on specific sets of data than many of the other models tested. While these two models still exhibit large errors, they provide the best fit of those tested and potentially suggest that there are uniform scaling realtionships between certain hydraulic properties and $k_{o_2}$. @wangPhysicallyBasedScaling2021's better fit, and more parsimonious, of the two models is reprinted as equation 3 (with that paper's reported coefficients of determination $r^2$). $\alpha$ is the fitted parameter in the linear regression. Equation 3-2nd line was fit to 588 field measurements of $k_{o_2}$ while equation 3-third line was fit to 3,919 simulated $k_{o_2}$ values at 35 rivers.
$$\mathbf{(3)} \begin{eqnarray} k_{o_2} &=& \alpha U_\star \\ &=& 47.67U_\star (r^2:0.53, measured) \\ &=& 49.71U_\star (r^2: 0.76, simulated)  \end{eqnarray}$$

**I'm struggling with this paragraph, and also don't know how to write it diplomatically**
We implement this model within BIKER (using 48 as a reasonable value for $\alpha$ given both forms of equation 3) for a few reasons. First, @wangPhysicallyBasedScaling2021 suggest that an $\alpha$ of approximately 48 is uniform across rivers and streams, regardless of the training data used. This is encouraging for use within BIKER as BIKER is specifically designed to be as river-agnostic as possible and so we sought as generalized an equation as possible. Second, it yields a simple linear relationship across all rivers, regardless of their size or steepeness. This is not necessairly true of other *k* models, for example those based on *eD* (section 1). The relationship between *eD* and *k* is not necessairly linear [@raymondScalingGasTransfer2012a], suggesting poorer performance for small *k* values. Further, recent work has shown that *k* does not scale with *eD* via the same statistical parameters across all river geomorphologies [@ulsethDistinctAirWater2019a]. We further explored the relationship presentedby [@ulsethDistinctAirWater2019a] and found that the $k \sim eD$ relationship fundamentally breaks down in rivers with low *eD*, which corresponds to nearly all rivers that SWOT will observe (Figure S2). Therefore, an *eD* based predictive model is less useful in SWOT rivers. Third, we sought to explicitly predict $k_{o_2}$, as its use extends beyond constraining gas fluxes from rivers and into parameterizing river metabolism models (equation 2). We stress that most other predictive models for *k* could be implemented within BIKER. However, our goal of using BIKER on any river that is observable by SWOT necessitates that we pick an equation that is applicable across as many rivers as possible. Future work should look at using a river-specific *k* equation to pseudo-calibrate BIKER to a specific river.

@@

BIKER, and Bayesian inference in general, starts from Bayes rule (equation 4), where $\Theta$ is some set of non-remotely-sensible parameters we want to solve for (including $k_{o_2}$), *x* is the observed data, $f(x|\Theta)$ is the 'likelihood function' or sampling model conditional on the parameters, and $p(\Theta)$ is the joint prior distribution of the parameters. Therefore, we are interested in solving for $p(\Theta|x)$, or the 'posterior' distribution. For BIKER, *x* is the SWOT-observables *W* and *H*. Note that $p(x)$ is usually computationally intractable to integrate, so Bayesian inference tools require only the proportionality to be specified: $p(\Theta|x) \propto f(x|\Theta)p(x)$. Sampling algorithms are then used to approximate the posterior distribution, as is done in BIKER and detailed below. First, we detail our likelihood specification and then our prior specifications.

@@

####3.2.1 BIKER likelihood
To concieve of equation 3 as a likelihood function, $k_{o_2}$ first needs to be written as a function of SWOT-observables *W* and *H* (which provides water surface slope $S_h$). This algebra is carried out and results in equation 5, where we assume that $R_h \approx \frac{A}{W}$ where *A* is wetted channel area. *A* is further split into the SWOT-observable portion *dA* and the unobservable portion $A_0$ following @durandEstimatingReachaveragedDischarge2014a. *dA* is estimated assuming a rectangular river channel so that $dA \approx \delta H*W$.

@@

Next, equation 5 is written as a Bayesian sampling model, in which all of the SWOT observations are sampled from the unknown model parameters ($A_0 + dA$ and $k_{o_2}$). This is equation 6. The $\sigma_{k_{o2}}^2$ parameter refers to the uncertainty inherent in equation 5's estimates. This will be explained in detail below. Note that in the formal model specification, equation 6 is written as a normal distribution of its log-transformed quantities. Here, it is written as a lognormal distribution for succinctness.

@@

####3.2.2 Prior Specifications
Equations 4 and 6 necessitate that we specify prior distributions for the parameters $A_0 + dA$ and $k_{o_2}$. Prior distributions formalize the a priori estimates (and uncertainties) for the non-remotely-sensed parameters. More intuitively, BIKER priors represent our 'prior river knowledge' of what $A_0 + dA$ and $k_{o_2}$ probably are for some river since they cannot be remotely sensed. All priors are formalized within the model as truncated normal distributions of the log-transformed terms such that $log(X) \sim N(\mu, \sigma^2)$ for $\lambda < log(X) < \epsilon$, using prior hyperparameters mean ($\mu$), standard deviation ($\sigma$), and upper ($\epsilon$) and lower bounds ($\lambda$) for any parameter *X*. In order to avoid relying on in situ information, as is the goal of BIKER and SWOT, we assign prior hyperparameters using SWOT data only. $A_0$ prior hyperparameters were assigned following the method developed by @brinkerhoffConstrainingRemoteRiver2020a. They developed a set of river channel prior hyperparameters for McFLI algorithms that are entirely RS-able and reflect differential channel hydraulics as a function of river geomorphology. They used an extensive database of field measurements and machine learning to identify patterns that associate river width with the hydraulic priors needed to run McFLIs so that prior hyperparameters may be assigned to rivers using only the existing remotely sensed data. This leaves the $k_{o_2}$ hyperparameters to be defined. We assigned those using a simple slope regression model trained on an available subset of the @raymondScalingGasTransfer2012a dataset of field-measured $k_{o_2}$: equation 7 assigns $\mu$ while $\sigma^2$ was set to equation 7's log-transformed standard error: 0.77. $\lambda$ and $\epsilon$ were set to log(0.0001) and log(60), respectively.
